The Minimal Retroreflective Microfacet Model

### 2025-12-23 article.tex
Typo fixes

%\paragraph*{Shadowing-Masking.}
%
%The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility %shadowing function
%$G_1(\vecl, \vech)$ for the light direction. In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of %\citet{HeitzMicrofacet} for a discussion of separable vs. non-separable shadowing-masking functions. Our derivation holds for any valid $G_2$ function, but we %focus here on the separable case for simplicity.}
%\begin{equation}
%G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \, G_1(\vecl, \vech)\ .
%\end{equation}
%Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
%\begin{equation}\label{eq:visible}
%  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
%\end{equation}
%Popular choices for $G_1$ with this property are the Smith masking function
%\begin{equation}
%  G_{1,\text{Smith}}(\vecv, \vech) = \frac{\chi^+(\vdot{\vecv}{\vech})  \avdot{\vecv}{\vecn}}{\int_{\Omega_{\vecv}} \cldot{\vecv}{\vech} D(\vech) \diff\vech},
%\end{equation}
%where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
%\begin{equation}
%  G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
%  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
%\end{equation}
%which is generally applicable for distributions with symmetry around the normal.

@@

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retroreflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector $\vecvv$ instead. From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
\begin{equation}
  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
\end{equation}
Thus for both reflection and refraction, the effect on the microsurface BSDF of using the back-vector as selector is to replace $\vecv$ by $\vecvv$ in the Fresnel term and in the denominators.

@@

The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry. It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction: \footnote{This is the seperable form commonly used in microfacet models; see Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs. non-separable shadowing-masking functions. In general, $G_{2}(\vecv, \vecl, \vech)$ is some function of $G_1(\vecv, \vech)$ and $G_1(\vecl, \vech)$, and thus reciprocity still holds.}
%In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs. non-separable shadowing-masking functions. Our derivation holds for any valid $G_2$ function, but we focus here on the separable case for simplicity.}
\begin{equation}
G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \, G_1(\vecl, \vech)\ .
\end{equation}
%Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
%\begin{equation}\label{eq:visible}
%  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
%\end{equation}
Popular choices for $G_1$
%with this property
are the Smith masking function
\begin{equation} \label{eq:smith_masking}
  G_{1,\text{Smith}}(\vecv, \vech) = \frac{\chi^+(\vdot{\vecv}{\vech})  \avdot{\vecv}{\vecn}}{\int_{\Omega_{\vecv}} \cldot{\vecv}{\vech} D(\vech) \diff\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
\begin{equation} \label{eq:vc_masking}
  G_{1,\text{VC}}(\vecv, \vech) =
  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

@@

The Fresnel term $F(\vecv, \vech)$ in the microfacet model accounts for the proportion of light reflected vs. refracted at the micro-surface. We have to verify that
\begin{equation}
F(\vecvv, \vech(\vecvv, \vecl)) = F(\vecll, \vech(\vecll, \vecv)) \ .
\end{equation}
Reflection symmetry of the Fresnel term implies
\begin{equation}
  F(\vecvv, \vech(\vecvv, \vecl)) = F(\vecv, \vech(\vecv, \vecll))
\end{equation}
and reciprocity symmetry further implies
\begin{equation}
  F(\vecv, \vech(\vecv, \vecll)) = F(\vecll, \vech(\vecll, \vecv)) \ .
\end{equation}
from which the desired equality follows.

>>>>

%\paragraph*{Shadowing-Masking.}
%
%The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility %shadowing function
%$G_1(\vecl, \vech)$ for the light direction. In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of %\citet{HeitzMicrofacet} for a discussion of separable vs. non-separable shadowing-masking functions. Our derivation holds for any valid $G_2$ function, but we %focus here on the separable case for simplicity.}
%\begin{equation}
%G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \, G_1(\vecl, \vech)\ .
%\end{equation}
%Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible microsurfaces \cite{HeitzMicrofacet}
%\begin{equation}\label{eq:visible}
%  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
%\end{equation}
%Popular choices for $G_1$ with this property are the Smith masking function
%\begin{equation}
%  G_{1,\text{Smith}}(\vecv, \vech) = \frac{\chi^+(\vdot{\vecv}{\vech})  \avdot{\vecv}{\vecn}}{\int_{\Omega_{\vecv}} \cldot{\vecv}{\vech} D(\vech) \diff\vech},
%\end{equation}
%where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
%\begin{equation}
%  G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
%  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
%\end{equation}
%which is generally applicable for distributions with symmetry around the normal.

@@

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retroreflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector $\vecvv$ instead. From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
Similarly, for the transmissive microsurface BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
\begin{equation}
  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
\end{equation}
Thus for both reflection and refraction, the effect on the microsurface BSDF of using the back-vector as selector is to replace $\vecv$ by $\vecvv$ in the Fresnel term and in the denominators.

@@

The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry. It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction: \footnote{This is the separable form commonly used in microfacet models; see Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs. non-separable shadowing-masking functions. In general, $G_{2}(\vecv, \vecl, \vech)$ is some function of $G_1(\vecv, \vech)$ and $G_1(\vecl, \vech)$, and thus reciprocity still holds.}
%In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs. non-separable shadowing-masking functions. Our derivation holds for any valid $G_2$ function, but we focus here on the separable case for simplicity.}
\begin{equation}
G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \, G_1(\vecl, \vech)\ .
\end{equation}
%Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible microsurfaces \cite{HeitzMicrofacet}
%\begin{equation}\label{eq:visible}
%  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
%\end{equation}
Popular choices for $G_1$
%with this property
are the Smith masking function
\begin{equation} \label{eq:smith_masking}
  G_{1,\text{Smith}}(\vecv, \vech) = \frac{\chi^+(\vdot{\vecv}{\vech})  \avdot{\vecv}{\vecn}}{\int_{\Omega_{\vecv}} \cldot{\vecv}{\vech} D(\vech) \diff\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
\begin{equation} \label{eq:vc_masking}
  G_{1,\text{VC}}(\vecv, \vech) =
  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

@@

The Fresnel term $F(\vecv, \vech)$ in the microfacet model accounts for the proportion of light reflected vs. refracted at the microsurface. We have to verify that
\begin{equation}
F(\vecvv, \vech(\vecvv, \vecl)) = F(\vecll, \vech(\vecll, \vecv)) \ .
\end{equation}
Reflection symmetry of the Fresnel term implies
\begin{equation}
  F(\vecvv, \vech(\vecvv, \vecl)) = F(\vecv, \vech(\vecv, \vecll))
\end{equation}
and reciprocity symmetry further implies
\begin{equation}
  F(\vecv, \vech(\vecv, \vecll)) = F(\vecll, \vech(\vecll, \vecv)) \ .
\end{equation}
from which the desired equality follows.
